### 2022-02-03 paper/lipics-v2021-sample-article.tex
init fscd draft

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{A Simple Type Theory for Category Theory} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Max S. New}
       {University of Michigan Computer Science \& Engineering, Ann Arbor, Michigan, USA \and \url{http://www.maxsnew.com} }
       {maxsnew@umich.edu}
       {}
       {}
       %% {https://orcid.org/0000-0002-1825-0097}
       %% {(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name
       %% ; only 1 author per \author macro; first two parameters are mandatory, other parameters ca
       %% n be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Daniel R. Licata}{Wesleyan University, Middletown, Connecticut, USA}
       {dlicata@wesleyan.edu}{}{}

\authorrunning{M.S. New and D.R. Licata} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Max S. New and Daniel R. Licata} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%% \acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
  We present a domain-specific type theory for formal category theory
  constructions and proofs. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to a form of ordred
  logic, we guarantee that all mappings between categories are
  functorial and all transformations are natural by construction, with
  no separate proof necessary. We demonstrate the utility of the type
  theory by proving a form of the Yoneda lemma which reads as the
  classical proof, but with all naturality side-conditions
  automatically satisfied.

  We show that the type theory is sound and complete for virtual
  equipments, a kind of double category which is known to model not
  just ordinary categories but enriched and internal categories as
  well. This means that proofs and constructions in our theory hold
  not just for ordinary categories but also for a broad class of
  generalized categories. We demonstrate that the type theory
  simplifies working with virtual equipments compared to their
  algebraic description in the literature.
\end{abstract}

\section{Introduction}

Category theory is a branch of mathematics that formalizes the
algebraic structure of function composition. The high level of
abstraction allows for succinct proofs relevant to structures across a
broad variety of mathematical domains. Category theory is of
particular interest in the theory of programming languages and proof
assistants, where it is commonly used to aid in the design of typed
languages, proof assistants and functional programming abstractions.

Novices in category theory are often overwhelmed by the number of
naturality and functoriality side conditions that abound in otherwise
simple proofs in category theory. For instance consider the proof of
the Yoneda lemma, often the first ``big'' proof in an introductory
category theory text. ... Here the reader is left to themselves to
show that the ``evaluation at identity'' morphism and its inverse are
actually natural transformations. They will then work through a
diagram chase to determine that this naturality condition indeed
holds.

Experienced category theorists have no trouble with this, as they are
familiar with a great many principles for constructing transformations
that are guaranteed to be natural. Consider an analogy with
calculus. When a mathematician writes a formula such as

$$\sin(f(x) - g(x))$$

They do not need to spend time with an explicit $\epsilon,\delta$
proof that the function is continuous or furthermore smooth. Instead,
they follow syntactic principles that compositions of continuous
functions are continuous.

There are similar syntactic notions for guaranteeing that
transformations are natural.

\subsection{Profunctors}

The syntax of Judgment Theory gives not just a syntax for the familiar
trio of category, functor and natural transformation, but additionally
a concept that is not usually highlighted in introductory category
theory texts: the profunctor. A simple intuition is that profunctors
are to functors as relations are to functions. Indeed, the syntax for
judgment theory has a straightforward semantics in the category of
sets and relations.

The reason profunctors are not commonly introduced is that they are
seen as an auxiliary notion: a profunctor $R : C \relto D$ is
typically defined as a functor $C^{\textrm{op}}\times D \to
\textrm{Set}$. However this definition requires two concepts that are
somewhat problematic: the opposite category $C^\textrm{op}$ and the
category of (small) sets $\textrm{Set}$. However, these notions do not
necessarily exist in all of our intended applications. First, we are
interested in semantics where are objects are \emph{multicategories}
where morphisms have multiple inputs but only a single output. In this
case the opposite of a multicategory is not a multicategory at all but
instead some kind of ``co-multicategory'' with multiple outputs but
one input. Second, having a universe object like $\textrm{Set}$ is also

\subsection{Applications to Enriched, Internal and Multicategory Theory}

The syntax of judgment theory is an internal language for
\emph{virtual equipments}. These are a kind of joint generalization of
double categories and multicategories.

Notably, the definition of a category in \secref{sec:internal} can be
applied as an extremely general notion of category, including internal
categories and enriched categories. This general notion is often
called a ``generalized multi-category'' and encompasses an astonishing
range of mathematical structures, including categories, topological
spaces, metric spaces, etc.

\section{Syntax of Judgment Theory}

\subsection{Positive Multiplicatives}

\subsection{Negative Multiplicatives}

\subsection{Additives}

\section{Synthetic Category Theory}

\section{Internal Category Theory}

\subsection{Duality}

\section{Semantics in Virtual Equipments}

\section{Related Work}

Cartesian Bicategories.

Directed Type theory

Fibrational approaches to logical relations

Double-cat-like syntax: Gradual Type Theory, System P of parametricity, cubical stuff?

String Diagrams?

%% \section{Typesetting instructions -- Summary}
%% \label{sec:typesetting-summary}

%% LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
%% In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
%% Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others. The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.

%% Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
%% \paragraph*{Minimum requirements}

%% \begin{itemize}
%% \item Use pdflatex and an up-to-date \LaTeX{} system.
%% \item Use further \LaTeX{} packages and custom made macros carefully and only if required.
%% \item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
%% \item Provide suitable graphics of at least 300dpi (preferably in PDF format).
%% \item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
%% \item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
%% \item Use a spellchecker to correct typos.
%% \end{itemize}

%% \paragraph*{Mandatory metadata macros}
%% Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
%% Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
%% \begin{itemize}
%% \item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
%% \item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
%% \item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
%% The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
%% \item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
%% \item \verb+\keywords+ (a comma-separated list of keywords).
%% \item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
%% \item \verb+\begin{abstract}...\end{abstract}+ .
%% \end{itemize}

%% \paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
%% Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults. To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
%% In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}. In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
%% \begin{itemize}
%% \item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
%% \item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
%% \item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
%%  \verb+\begin{enumerate}[(a)]+ or the like.)
%% \item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
%% \item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
%% \item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
%% \item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
%% \item Do not change the bibliography style. In particular, do not use author-year citations. (The
%% \texttt{natbib} package is not supported.)
%% \end{itemize}

%% \enlargethispage{\baselineskip}

%% This is only a summary containing the most relevant details. Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
%% \href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}

%% \section{Lorem ipsum dolor sit amet}

%% Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

%% \begin{lemma}[Lorem ipsum]
%% \label{lemma:lorem}
%% Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
%% \end{lemma}

%% \begin{proof}
%% Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.

%% \proofsubparagraph*{Just some paragraph within the proof.}
%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
%% \begin{claim}
%% content...
%% \end{claim}
%% \begin{claimproof}
%% content...
%%     \begin{enumerate}
%%         \item abc abc abc \claimqedhere{}
%%     \end{enumerate}
%% \end{claimproof}

%% \end{proof}

%% \begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
%% \label{lemma:curabitur}
%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
%% \end{corollary}

%% \begin{proposition}\label{prop1}
%% This is a proposition
%% \end{proposition}

%% \autoref{prop1} and \cref{prop1} \ldots

%% \subsection{Curabitur dictum felis id sapien}

%% Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

%% \begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
%% for i:=maxint to 0 do 
%% begin 
%%     j:=square(root(i));
%% end;
%% \end{lstlisting}

%% \subsection{Proin ac fermentum augue}

%% Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

%% \begin{itemize}
%% \item Ut vitae diam augue. 
%% \item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
%% \item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
%% \end{itemize}

%% \begin{remark}
%% content...
%% \end{remark}

%% \section{Pellentesque quis tortor}

%% Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

%% \begin{lemma} [Quisque blandit tempus nunc]
%% Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
%% \end{lemma}

%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

%% \section{Morbi eros magna}

%% Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%% %%
%% %% Bibliography
%% %%

%% %% Please use bibtex,

\bibliography{lipics-v2021-sample-article}

\appendix

\section{Additional Rules}

%% \section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

%% List of different predefined enumeration styles:

%% \begin{itemize}
%% \item \verb|\begin{itemize}...\end{itemize}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{itemize}

%% \begin{enumerate}
%% \item \verb|\begin{enumerate}...\end{enumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{enumerate}

%% \begin{alphaenumerate}
%% \item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{alphaenumerate}

%% \begin{romanenumerate}
%% \item \verb|\begin{romanenumerate}...\end{romanenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{romanenumerate}

%% \begin{bracketenumerate}
%% \item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{bracketenumerate}

%% \begin{description}
%% \item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
%% \item[Description 2] Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \item[Description 3]  \dots
%% %\item \dots
%% \end{description}

%% \cref{testenv-proposition} and \autoref{testenv-proposition} ...

%% \section{Theorem-like environments}\label{sec:theorem-environments}

%% List of different predefined enumeration styles:

%% \begin{theorem}\label{testenv-theorem}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{theorem}

%% \begin{lemma}\label{testenv-lemma}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{lemma}

%% \begin{corollary}\label{testenv-corollary}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{corollary}

%% \begin{proposition}\label{testenv-proposition}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{proposition}

%% \begin{conjecture}\label{testenv-conjecture}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{conjecture}

%% \begin{observation}\label{testenv-observation}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{observation}

%% \begin{exercise}\label{testenv-exercise}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{exercise}

%% \begin{definition}\label{testenv-definition}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{definition}

%% \begin{example}\label{testenv-example}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{example}

%% \begin{note}\label{testenv-note}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{note}

%% \begin{note*}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{note*}

%% \begin{remark}\label{testenv-remark}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{remark}

%% \begin{remark*}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{remark*}

%% \begin{claim}\label{testenv-claim}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claim}

%% \begin{claim*}\label{testenv-claim2}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claim*}

%% \begin{proof}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{proof}

%% \begin{claimproof}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claimproof}

\end{document}