### 2022-12-08 paper/reviews/response.txt
reformatted response

We thank the reviewers for their very detailed and helpful feedback.  We
will integrate them into the final version of the paper --- including
all of the clarification comments not mentioned below.

We respond to the most pertinent questions first. And each response
includes "Reviewer X" so you can search for the responses to your
specific questions.

We agree with Review 3 that the most important advantages of our
approach are for enriched category theory (indeed, this was our
motivation for starting this work), and can revise the introduction to
emphasize this aspect more.  However, we do feel there are some
advantages even for ordinary category theory: while a synthetic theory
cannot eliminate all side conditions, it can help structure when the
user really needs to prove them.  One of the examples the reviewer
mentions, respect for quotients, is a classic example where an
"analytic" approach is unpleasant in practice ("setoid hell"), and
improved by synthetic approaches in NuPRL/Observational Type
Theory/HoTT/Cubical type theory.  While an adept category theorist may
know how to manage these obligations and e.g. pick a definition that
is more automatically natural, our experience explaining category
theory to programming languages researchers as background for
understanding some of our past papers suggests that the support of a
type theory may be useful for explaining how experts think about these
issues. We agree with the reviewer that the form the "domain specific
proof assistant" could take is as a library design for an existing
proof assistant, i.e., an embedded DSL.

Review 3 asks what the relevant CwF structures are in Constructions 4
and 5.  Our type theory distinguishes between the virtual equipment
layer (categories, functors, profunctors, natural transformations) and
the "meta" layer, a dependent type theory.  The semantics interprets
this in a hyperdoctrine of virtual equipments, with the CwF structure
for the meta-layer on the base of the hyperdoctrine.  In the "standard"
model, the base is just the category of sets Set, while the fibers of
the hyperdoctrine are Cat.  So there is no need for a CwF structure on
Cat.

Review 3 asks if every CwF gives rise to a HVE. Certainly construction
4 can be adapted to any elementary topos with sufficient universes
since the category-theoretic constructions needed are all
intuitionistically valid. So this should correspond to a Category with
Families extended with certain universal properties, for instance,
extensional identity types.

Review 3 suggests some additional references and examples (Palmgren,
  Weber, Di Liberti and Loregian). Thank you; we will include them!

Review 1 asks about the limitations of the present formalism, and
Review 2 similarly asked about why we restrict to unary
functors/binary profunctors. There are several useful concepts we
cannot formalize because our judgments are limited to covariant
functors of a single argument: categories of functors and opposite
categories, for instance. Next steps for the current work would be to
extend the judgmental structure to support these. Some preliminary
work along these lines is in Reference [21]. However, there is a
tradeoff: by adding more structure in the type theory, we support
fewer models. We chose these restrictions as a starting point because
they correspond to a well-understood notion of model (virtual
equipments), future extensions would likely require the equipment to
come with a monoidal closed structure. In particular, we expect that
categories enriched in a non-symmetric monoidal closed category will
not be a model of these extensions.

Review 1 asks about the definition of d(Ξ) and the presheaf
elimination rules. d-(x:C) and d+(x:C) are both x:C; we will add this
definition. The negative presheaf elimination rule is the closure
under substitution of a : C ; p : P- C |- a in p Set i.e. evaluation
is a map C^op x (Set^{C^op}) -> Set

Review 1 asks about the tensor product of profunctors. The tensor
product of profunctors is indeed a colimit --- e.g. it can be defined
using a coend.  Syntactically, beta is just a bound variable (and can
of course be alpha-converted).  The fact that maps out of the tensor
respect the quotienting is implicit/synthetic, and comes from the fact
that all open terms ... beta : C ... |- t : S respect this
quotienting.  This is similar to how quotient types are handled in
NuPRL/Observational type theory/HoTT/Cubical type theory.

Review 1 had several questions about Section 4. We will rephrase
Theorem 2 with less shorthand. The formal meaning of "almost any" is
to emphasize that the initiality theorem is modular in that most
connectives are independent of each other. The exception is that to
get the right universal property for hom and tensor of profunctors, we
need the exponential of profunctors. vDbl_r and Cat_U should, as the
reviewer guessed, be vEq and Cat_κ. The purpose of Construction 3 is
to reduce the construction of a model of VETT to just constructing a
virtual equipment, since any virtual equipment can be "upgraded" to a
hyperdoctrine of equipments using this construction.

Review 2 asks if one can generate explicit proofs of functoriality
/naturality from proofs in the type theory.  Yes, "running" the
(constructive) proof of soundness (Theorem 2) would do exactly this, by
compiling our "synthetic" proofs to "analytic" ones.

Review 2 asks about "P ▹∀α Q" and the difference with "∀α.P".

P ▹∀α Q is the 'hom' of profunctors, the right adjoint to - (x)∃α P.  This
type binds both an object variable α and a transformation variable x:P.

In the type of composition 
   ∀α1. (α1 → α2) |>∀α2 (α2 → α3 |>∀α3 (α1 → α3))
α2 occurs covariantly in (α1 → α2)
and covariantly in (α2 → α3 |>∀α3 (α1 → α3)) because the
contravariance of |> and → cancel out.

∀α.P is the 'end' over a profunctor with 1 free variable (Figure 5,
bottom row).  This type is used to bind the "last" object variable in a
singleton context. The type for functoriality that the reviewer gives is
isomorphic to the type we gave by Lemma 3, part 5.

Review 2 asks about our comments on the Yoneda lemma. We had a typo in
this formula, thanks --- we meant ∀x. [ (∀y.(x=y) ⇒ Py) iff P x].

Review 2 asks about the proof of initiality theorem. The initiality
theory could in principle be mechanized in Agda, though we have not
yet done so.